Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-04 android-stacks.tex
results - question 3

%the mobile industry shipped over 315 million smartphones in the last quarter of 2013
%The pervasiveness of the World Wide Web and the astonishing increase 
%in the number of smartphones stimulated the creation of applications which extend the phones
% capabilities far beyond of the basic calls and textual messages. 
%f we look at global smartphone shipments the numbers are astonishng;
%the mobile industry shipped over 315 million smartphones in the last quarter of 2013 [ref-googleIO].
%Such increase on smartphones combined with the pervasiveness of the World Wide Web
%have stimulated the creation of applications which extend the phones caVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE QUEBRADO... PELO EXCEPTION Mpabilities 
%far beyond of the basic calls and textual messages. 
315,92
In this context, Android has become a leading platform - thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
 own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
and the number of apps built on top of Android increases in a daily rate.

%ref: http://www.gartner.com/newsroom/id/2645115
% Gartner, Inc. Press release, 2012. www.gartner.com/newsroom/id/2237315.
% http://www.gartner.com/newsroom/id/2645115

%If on the one hand such applications are getting mode and more popular....
%On the other hand... there is a need for reliability, as some core functionalities 
%are moving to apps....such applications impose more challenges to software reliability 
%Such mobile platforms are intrinsically multi-threaded and support a development model 
% slightly (or sometimes drastically) different from those adopted on Web development [arie-icse],  
%abnormal computation states

%In the same rate the number of a users increase, also increases 
%the number of users affected by application faults and crashes.

@@

%Hence, in the same rate the number of mobile users increase, also increases 
%the number of users affected by such threats.
%faulty apps and app crashes.

@@

% Google estimates that Android has over 1 billion active users per month, 
% who pull out theirs phones and check it over 100 billion times.

%If we google for "Android app crashes"
% we can find 1.200.000 pages about the subject on forums and blogs. In Feb. 2014
%Gartner issue a report on mobile technologies and capabilities 
%[ref-https://www.gartner.com/doc/2665315/top--mobile-technologies-capabilities], 
%which emphasized the importance of monitoring application crashes.

Our conjecture is that several Android applications may be suffering from common 
fragilities on the exception-related code (of the applications themselves or 
the underlaying platform) which may be degratind the robustness of such applications.
We call ``fragilities" characteristics on the exception-related code that favor the introduction
of aforementioned failures.

%, and that such fragilities may be leading to
%decrease the application robustness, leading for instance to the Uncaught Exception problem -
% which is one of the main causes of application crashes ~\cite{jo2004uncaught, zhang2012amplifying}.
%investigate the whether the exception stack traces
% can reveal common characteristics on the exception-related code that
%may be favoring the introduction of the aforementioned failures, which would be responsible for 
%decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.

%[ref AST].

@@

% The
%concentration of open source projects and the existence of initiatives that mine
%such information, such as GHTorrent~\cite{Gousi13}, allow the information about
%projects to be visible within and across projects and make a full population
%studies possible.

%Among the issues defined on issue trackers, we can find issues reporting bugs, and system crashes 
%found as a result of application executions. Such issues usually contain exception stack traces [ref][ref]. 
%<SUCH ISSUES INCLUDE STACK TRACES. WHAT IS THE RELATION BETWEEN STACK TRACE AND CRASH>
%<stack traces are part of good issues....>

%Stack traces are a useful source of information about system crashes, and is often used to support developers in debugging~\cite{schroter2010stack}. Moreover, information %from stack traces 
%combined with crash and bug reports can be used to support bug classification and clustering~\cite{wang2013improving, kim2011crash, %dhaliwal2011classifying}, fault-proneness prediction %models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.

@@

%Most of such studies, however, focused the analysis on one or two system at a time.
%This paper reports an exploratory study which targets a family of applications 
%(sharing the common Android framework) aiming at extracting 
%the common characteristics of exception stack traces reported to such systems to provide
%insights on similar fragilities on the exception related code of the applications and underlaying framework.

%482 applications) and Google code (157 applications)
%GitHub 31,592
% Google Code: 127,456
%All: 159,048

@@

%projects: 788	 - projects with issues - 724 - 
%projects with issues (with exceptions): 243 
%projects with stacks: 158  -  stacks: 3559 (defect)
%Github: issues: 50342 - stacks: 3865 - projects: 2542 - projects with stacks: 293

%To guide our analysis, we compiled a set of best practices on how to use
%exceptions by composing relevant suggestions by Gosling~\cite{gosling2000java},
%Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.  The
%comparison of those with the mined exceptions revealed several interesting
%findings, such as:

@@

\emph{II-Error represents an unrecoverable condition detected by the runtime enviroment which
should not he handled} (\cite{gosling2000java}). It results from failures detected
by the Java Virtual Machine which indicate resource deficiencies, invariant
failures or other conditions that make impossible the program to recover.

@@

\subsection{What can the Exception Wrappings tell us?}
%\subsection{Exploring the Information Embedded on Exception Wrappings}

Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors.

@@

  \hline
      %none   & Runtime   &  2360 & 359 \\
      %none  &   Checked  & 422  & 117 \\
      %none  &   Error   & 381  & 141 \\
      %none &  Undefined  &  15   & 10  \\
      %none & Throwable  &  1    & 1 \\
     %\hline
      %Runtime   & Runtime & 560  & 182 \\    
      %Checked   & Checked & 98   & 42  \\
      %Error     & Error   & 15   & 14  \\
    %\hline
  \end{tabular}
\caption{Wrappings comprising different exception types.}
\label{tab:wrappingandroid}
\end{table*}

This analysis revealed unexpected wrappings such as: Error wrapping a Checked 
exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
Java is the only language that provides a hybrid exception model 
which offers  different types to represent different exception behaviors (i.e., error, 
runtime and checked) (see Section~\ref{sec:best}). For instance,
according to Java specification, Errors should not be handled inside the system 
since they usually represent unrecoverable conditions
detected by the runtime environment (e.g.,  OutOfMemoryError). On the other hand, checked exceptions should
represent recoverable conditions. Next we discuss and exemplify the cross-type wrappings
 found in this study.

\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
detected on 88 projects. Inspecting the elements responsible for performing such wrapping, we could observe that
approximately  50\% of such wrappings were performed on methods defined on Android
Platform.

Taking a closer look on the methods responsible for such wrappings,
 we could observe that approximately 41\% of them were performed by a single class,
 named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
(window) of an Android app.
Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
As we can see from lines 4-8, in case any exception is thrown in the context of such try-catch block, 
it is wrapped in a general Runtime exception and re-thrown.

\begin{table}
\centering
\begin{tabular}{lll}
    \hline
 \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
    \hline
android.app.ActivityThread & java.lang.RuntimeException & 61\\
android.view.LayoutInflater & android.view.InflateException &  7 \\
android.os.AsyncTask & java.lang.RuntimeException &  7\\
\hline
  \end{tabular}
\caption{Top 3 Runtime-Cheked wrappers.}
\label{tab:wrapping01}
\end{table}

%MOVED TO THE PICTURE
%{\footnotesize
%\begin{verbatim}
%1. private Activity performLaunchActivity(...) {
%2.  try {
%3.   ...
%4.   ...callActivityOnCreate(...);
%5.  } catch (Exception e) {
%6.    ...
%7.   throw new RuntimeException(..., e);
%8.  }
%9.}
%\end{verbatim}
%}

%SELECT SIGNALER_method, COUNT(*) AS N
%  FROM stacktracesummary WHERE ROOT_EXTENDS='CHECKED' AND EXCEPTION_EXTENDS='RUNTIME' and
%  ( SIGNALER_method like '%onDestroy%' or
%   SIGNALER_method like '%onCreate%' or
% SIGNALER_method like '%onStart%' or
% SIGNALER_method like '%onResume%' or
% SIGNALER_method like '%onPause%' or
% SIGNALER_method like '%onStop%' or
% SIGNALER_method like '%onDestroy%' )
%   GROUP BY SIGNALER_method
%  ORDER BY N DESC;

%SELECT COUNT(*) AS N
 % FROM stacktracesummary WHERE ROOT_EXTENDS='CHECKED' AND EXCEPTION_EXTENDS='RUNTIME' and
 % ( stack like '%onDestroy%' or
 %  stack like '%onCreate%' or
% stack like '%onStart%' or
% stack like '%onResume%' or
% stack like '%onPause%' or
% stack like '%onStop%' or
% stack like '%onDestroy%' );

%GC = 6

%However, since the Activity life clycle method (i.e., callActivityOnCreate(...)) that is invoked in this block,
% do not include checked exceptions on its signatures. Hence they could not throw a checked exception to be handled 
%in the method illustrated above. However the code inspection performed on such wrappings revelaed an
%``undocumented checked exception" signaled by this life cycle method.

%\emph{\textbf{Undocumented Checked Exceptions.}} There are usually
%2 possible for a checked exception to be thrown by a method:
%it can either be specified on the method signature or it can be wrapped in a 
%runtime exception and re-thrown.

@@

% Several types of exceptions were  wrapped the RuntimeExceptions thrown by the 
% ActivityThread. Table X illustrates them.
% #\begin{table}
% \centering
% \begin{tabular}{lll}
%    \hline
%    \bfseries{wrapper} & \bfseries{root} & \bfseries{\#} \\
%    \hline
% java.lang.RuntimeException & java.lang.ClassNotFoundException &	32 \\
% java.lang.RuntimeException & java.lang.NoSuchMethodException &	6 \\
% java.lang.RuntimeException & java.lang.InstantiationException) & 5 \\
% ...andlyticsproject.NetworkException & org.apache...HttpResponseException &	5\\
% android.view.InflateException & java.lang.NoSuchMethodException & 4 \\
% \hline
%  \end{tabular}
% \caption{Elements responsible for the wrapping.}
% \label{tab:wrapping01}
% \end{table}

% \begin{table}
% \centering
% \begin{tabular}{lll}
%   \hline
%   \bfseries{Exception} & \bfseries{Ocurrences} \\
%   \hline
% java.lang.RuntimeException &	69 & 61\% \\
% android.view.InflateException	& 10 & 8,7\% \\
% com.github.andlyticsproject.console.NetworkException	& 7 & - \\
% java.lang.IllegalStateException	& 6 & - \\
% \hline
%  \end{tabular}
% \caption{Elements tresponsible for the wrapping.}
% \label{tab:wrapping01}
% \end{table}

\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
on 46 projects on which an Error was wrapped in a runtime exception.
Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any throwable signaled in the context of an asynchronous 
task is converted in to a runtime and re-thrown. Some of the Erros wrapped on instances of runtime exceptions were 
java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}.

%ASSNCTASKS NON ANNOTATED METHODS…….
%SOME HAVE EXCEPTION INTERFACE 
%EXCEPTIONS ON MEHTODS SIGNATURE OTHERS NOT
%INCONSISTENCY…. DEPENDS ON DEVELOPERS 
%WORK INSTEAD OF TOOLS.

%{\footnotesize
%\begin{verbatim}
%protected void done() {
% ...
 %  try {
 %     ...
  % } catch (InterruptedException e) {
  %    android.util.Log.w(..., e);
  % } catch (ExecutionException e) {
  %    throw new RuntimeException("...",e.getCause());
  % } catch (CancellationException e) {
  %    ...
 %  } catch (Throwable t) {
 %    throw new RuntimeException("...", t);
 %  }
 %   ...
% }
%}
%\end{verbatim}
%}

@@

%\begin{table}
%\centering
%\begin{tabular}{ll}
% \bfseries{Wrapper type - Root cause type} & \bfseries{\#}\\
% \hline
%...android.robospice.CacheSavingException - java.lang.ClassCastException & 6 \\
%java.lang.reflect.InvocationTargetException - java.lang.NullPointerException &  3 \\
%java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException & 2 \\
%java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException & 2 \\
%  \end{tabular}
%\caption{Elements tresponsible for the wrapping.}
%\label{tab:wrapping01}
%\end{table}

\emph{\textbf{Checked Exception wrapping Error.}} Almost all of these wrappings
 were also caused by the reflection library 
used by Android applications. The methods responsible for the wrappings
were also native methods written in C. Table ~\ref{tab:exampleswrap} illustrates
some oth these wrappings some of them are masking an OutOfMemoryError
into a checked exception.

%\begin{table}
%\centering
%\begin{tabular}{lll}
% \bfseries{Wrapper type - Root cause type} & \bfseries{\#}\\
% \hline
%java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError & 5 \\
%java.lang.Exception - java.lang.OutOfMemoryError & 1 \\
%java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError &	1 \\
%java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError&	1 \\
%  \end{tabular}
%\caption{Elements tresponsible for the wrapping.}
%\label{tab:wrapcheckerr}
%\end{table}

\emph{\textbf{Error wrapping Checked Exception.}} Some of these wrappings were performed by methods defined by java.lang.Class methods such as 
at java.lang.Class.getDeclaredFields(Native Method)
at java.lang.Class.getDeclaredConstructors(Native Method), 
while others were performed by javax.crypto package. Table X illustrates
the exception types involved in such wrappings.

%\begin{table}
%\centering
%\begin{tabular}{lll}
% \bfseries{Wrapper} & \bfseries{Root} & \bfseries{\#}\\
% \hline
%java.lang.NoClassDefFoundError & java.lang.ClassNotFoundException & 4 \\
%java.lang.AssertionError & javax.crypto.ShortBufferException) & 3 \\
% \end{tabular}
%\caption{Elements tresponsible for the wrapping.}
%\label{tab:wrapping01}
%\end{table}

@@

%\begin{table}
%\centering
%\begin{tabular}{ll}
% \bfseries{Wrapper type - Root cause type} & \bfseries{\#} \\
% \hline
%java.lang.ExceptionInInitializerError - java.lang.NullPointerException & 9 \\
%java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException &	2 \\
% \end{tabular}
%\caption{Elements tresponsible for the wrapping.}
%\label{tab:wrapping01}
%\end{table}

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError).
The study also detected an undocumented checked exception signaled by
  a native method defined in Android.
}
\end{minipage}}

@@

\subsection{On the need of a rescue to the almost infeasible task of preventing uncaught exceptions}
In this study we could observe undocumented runtime exceptions thrown by third party code,
and even undocumented checked exception thrown by Java supporting library.
Such undocumented exceptions make difficult, and most of the times infeasible
for the client code to protect against `unforeseen" situations that may happen 
while calling a library code.

@@

\subsection{The exception handling confusion problem}
When (mis)applied, the exception wrapping can make the exception-related code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to an unpredictable state in the presence of exceptions. 
To illustrate this problem, we can use one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the runtime environment which make impossible the program to recover. 
Currently there is no way of enforcing Java exception type conventions during program development.
Hence, further investigation is needed on finding ways to help developers in dealing with such
 problem, either preventing odd wrappings or enabling the developer to better deal with them or even
researches on the real usefulness of Java hybrid exception model.

@@

\subsection{On the null pointer problem}

IThe null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake” ~\cite{hoare}:

In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speach.
This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations ~\footnote{The @Nullable annotations are used by tools such as FindBugs, Eclipse, NetBeans, IntelliJ to detect potential 
null pointer dereferences at compile time. In April/2014 the Android Studio 0.5.5 was 
released with support to these annotations};
(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
or (iii) whether language designs which avoid null pointers, such 
as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs. 
.

>>>>

In this context, Android has become a leading platform - thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
 own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
and the number of apps built on top of Android increases in a daily rate.

@@

@@

Our conjecture is that several Android applications may be suffering from common 
fragilities on the exception-related code (of the applications themselves or 
the underlaying platform) which may be degratind their robustness.
We call ``fragilities" characteristics on the exception-related code that favor the introduction
of aforementioned failures.

@@

@@

@@

@@

\emph{II-Error represents an unrecoverable conditions which should not he handled} 
(\cite{gosling2000java}).  Errors should result from failures detected
by the runtime environemt which indicate resource deficiencies, invariant
failures or other conditions, that make impossible for the program to recover.

@@

\subsection{What can the Exception Wrappings tell us?}

@@

  \hline
      %none   & Runtime   &  2360 & 359 \\
      %none  &   Checked  & 422  & 117 \\
      %none  &   Error   & 381  & 141 \\
      %none &  Undefined  &  15   & 10  \\
      %none & Throwable  &  1    & 1 \\
     %\hline
      %Runtime   & Runtime & 560  & 182 \\    
      %Checked   & Checked & 98   & 42  \\
      %Error     & Error   & 15   & 14  \\
    %\hline
  \end{tabular}
\caption{Wrappings comprising different exception types.}
\label{tab:wrappingandroid}
\end{table*}
Java is the only language that provides a hybrid exception model 
which offers three kinds of exceptions each one holding an inttended exception behavior (i.e., error, 
runtime and checked). The best practices I and II states the behavior that Error an Checked exceptions hold, which are: 
(i) Errors should not be handled inside the system since they usually represent unrecoverable conditions detected 
by the runtime environment (e.g.,  OutOfMemoryError); (ii) checked exceptions, on the other hand, should represent recoverable conditions.

Table~\ref{tab:wrappingandroid} presents all wrappings found in this study, that
envolve different exception types (i.e., Error, Checked, Exception). We call 
such a wrapping a ``cross-type wrapping".  Such wrappings reveals 
violations to best practices I and II (e.g. Error wrapping a Checked exception),
To illustrate the problem behind such wrappings, we can use one of the wrappings found in this study. 
When the developer is confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the runtime environment which make impossible the program to recover. 
Hence, trying to handle this exception may lead the program to an unpredictable state.

Next we discuss and exemplify the cross-type wrappings
 found in this study.

\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
detected on 88 projects. Inspecting the elements responsible for performing such wrapping, we could observe that
approximately  50\% of such wrappings were performed on methods defined on Android
Platform. Taking a closer look on the methods responsible for such wrappings,
 we could observe that approximately 41\% of them were performed by a single class,
 named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
(window) of an Android app.
Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
As we can see from lines 4-8, in case any exception is thrown in the context of such try-catch block, 
it is wrapped in a general Runtime exception and re-thrown.

%\begin{table}
%\centering
%\begin{tabular}{lll}
%   \hline
% \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
%    \hline
%android.app.ActivityThread & java.lang.RuntimeException & 61\\
%android.view.LayoutInflater & android.view.InflateException &  7 \\
%android.os.AsyncTask & java.lang.RuntimeException &  7\\
%\hline
%  \end{tabular}
%\caption{Top 3 Runtime-Cheked wrappers.}
%\label{tab:wrapping01}
%\end{table}

@@

\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
on 46 projects on which an Error was wrapped in a runtime exception.
Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any instance of Throwable signaled in the context of an asynchronous 
task is converted in to a runtime and re-thrown.  Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
Such wrapping masks the unrecoverable exception in a general excepion - doing so  that can be mistakenly handled

@@

\emph{\textbf{Checked Exception wrapping Error.}} Almost all of these wrappings
 were also caused by the reflection library 
used by Android applications. The methods responsible for the wrappings
were also native methods written in C. Table~\ref{tab:exampleswrap} illustrates
some oth these wrappings some of them are masking an OutOfMemoryError
into a checked exception.

\emph{\textbf{Error wrapping Checked Exception.}} These wrappings were performed
either by native methods defined by java.lang.Class (e.g., getDeclaredFields, getDeclaredConstructors) 
or methods from javax.crypto package. Table~\ref{tab:exampleswrap} illustrates
the exception types involved in such wrappings.

@@

@@

AS WE COULD OBSERVE...

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
MUDARRRR Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError).
The study also detected an undocumented checked exception signaled by a native method defined in Android.
}
\end{minipage}}

@@

\subsection{On the need of rescues to the almost infeasible task of preventing uncaught exceptions}
In this study we could observe undocumented runtime exceptions thrown by third party code,
and even undocumented checked exception thrown by Java supporting library.
Such undocumented exceptions make difficult, and most of the times infeasible
for the client code to protect against `unforeseen" situations that may happen 
while calling a library code.

@@

\subsection{On the null pointer problem}
The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake” ~\cite{hoare}:

In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speach.
This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations ~\footnote{The @Nullable annotations are used by tools such as FindBugs, Eclipse, NetBeans, IntelliJ to detect potential 
null pointer dereferences at compile time. In April/2014 the Android Studio 0.5.5 was 
released with support to these annotations};
(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
or (iii) whether language designs which avoid null pointers, such 
as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

\subsection{The exception handling confusion problem}
When (mis)applied, the exception wrapping can make the exception-related code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to an unpredictable state in the presence of exceptions. 
To illustrate this problem, we can use one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the runtime environment which make impossible the program to recover.

Currently there is no way of enforcing Java exception type conventions during program development.
Hence, further investigation is needed on finding ways to help developers in dealing with such
 problem, either preventing odd wrappings or enabling the developer to better deal with them or even
researches on the real usefulness of Java hybrid exception model.

@@
